---
title: "Model Explorer — Dashboard Setup"
author: ""
date: "`r Sys.Date()`"
output: html_document
---

# Model Explorer Setup

This notebook configures the **Model Explorer** dashboard for your project.
It patches the config block in `generate-explorer-data.R` (in place) and
copies a clean `model-explorer.html` one level up from `dashboard/`.

## 1. Configure your project

Edit the chunk below to match your project layout.

```{r config}
# ── TARGET DIRECTORY ────────────────────────────────────────────────────────
# Automatically set to one level up from this dashboard/ folder.
# model-explorer.html will be placed here, next to the phase directories.
TARGET_DIR <- dirname(normalizePath(".", mustWork = TRUE))

# ── PHASE MAP ───────────────────────────────────────────────────────────────
# Keys = exact subdirectory names under TARGET_DIR
# Values = display labels shown as phase tabs in the dashboard
# Only directories listed here will be scanned for models.
# Example: if your layout is models/single_route/ and models/multiple_route/
#   then use: "single_route" = "Single Route", "multiple_route" = "Multiple Route"
PHASE_MAP <- list(
  # "directory_name" = "Display Label",
  # "01-Primary"     = "Primary",
  # "02-Decay"       = "Decay"
)

# ── NOTES EXCLUDE ───────────────────────────────────────────────────────────
# Parameters that are always fixed by design across all models.
# These are omitted from the "fixed params" notes in the Overview tab
# to reduce clutter. (They'll still appear in the Parameters tab.)
NOTES_EXCLUDE <- c(
  # "k_pop",
  # "rebound_flag_pop"
)
```

## 2. Expected directory layout

The dashboard expects Monolix projects organized like this:

```
final-models/
├── model-explorer.html          <- the dashboard (created by this notebook)
├── dashboard/
│   ├── generate-explorer-data.R <- data generator (config patched by this notebook)
│   ├── model-explorer-template.html <- clean HTML template
│   ├── setup-dashboard.Rmd      <- this file
│   ├── validate-models.R        <- validation checks
│   ├── wipe-data.R              <- strips injected data from HTML
│   └── update-dashboard.sh      <- shell wrapper for generate-explorer-data.R
│
├── 01-Primary/                  <- phase directory (name must match PHASE_MAP)
│   ├── model_family_A/          <- family grouping (sidebar subheader)
│   │   ├── model_name/          <- one per model
│   │   │   ├── model_name/      <- Monolix output directory
│   │   │   │   ├── populationParameters.txt
│   │   │   │   ├── logLikelihood.txt
│   │   │   │   ├── summary.txt
│   │   │   │   ├── IndividualParameters/
│   │   │   │   │   └── estimatedIndividualParameters.txt
│   │   │   │   └── ChartsData/
│   │   │   │       ├── IndividualFits/
│   │   │   │       └── ObservationsVsPredictions/
│   │   │   └── model_name.mlxtran
│   │   └── another_model/
│   │       └── ...
│   └── model_family_B/
│       └── ...
│
├── 02-Decay/
│   └── ...
```

Key conventions:

- **Phase directories** are top-level folders whose names appear as keys in `PHASE_MAP`
- **Family directories** sit one level below phases; models within the same family
  are grouped together in the sidebar under a shared header
- **Model directories** contain a Monolix output subdirectory with the same name,
  plus a `.mlxtran` project file alongside it
- The model output directory **must contain `populationParameters.txt`** — this is
  how the scanner identifies valid model runs

## 3. Deploy HTML

Copy the template HTML one level up from `dashboard/` so it sits next to
the phase directories (e.g., `final-models/model-explorer.html`).

```{r html, message=TRUE}
source_template <- "model-explorer-template.html"
if (!file.exists(source_template)) {
  stop("model-explorer-template.html not found next to this Rmd")
}

html <- readLines(source_template, warn = FALSE)

# Wipe any injected data so we start clean
start_idx <- grep("// <!-- MODEL_DATA_START -->", html, fixed = TRUE)
end_idx   <- grep("// <!-- MODEL_DATA_END -->", html, fixed = TRUE)

if (length(start_idx) > 0 && length(end_idx) > 0) {
  html <- c(
    html[1:start_idx],
    "window.MODEL_DATA = [];",
    "window.DASHBOARD_META = {};",
    html[end_idx:length(html)]
  )
}

target_html <- file.path(TARGET_DIR, "model-explorer.html")
writeLines(html, target_html)
cat("\u2713", target_html, "\n")
```

## 4. Patch generate-explorer-data.R config

Rewrites the config section of the existing `generate-explorer-data.R`
with your `PHASE_MAP`, `NOTES_EXCLUDE`, and `BASE_DIR`.

```{r patch-config, message=TRUE}
gen_path  <- "generate-explorer-data.R"
gen_lines <- readLines(gen_path, warn = FALSE)

# Find config section boundaries
config_start <- grep("^# ── Config", gen_lines)
config_end   <- grep("^# ── Helper", gen_lines)[1] - 1

if (length(config_start) == 0 || is.na(config_end)) {
  stop("Could not find config section markers in generate-explorer-data.R")
}

# Build the PHASE_MAP string
if (length(PHASE_MAP) > 0) {
  phase_entries <- paste0('  "', names(PHASE_MAP), '" = "', unlist(PHASE_MAP), '"')
  phase_str <- paste0("PHASE_MAP <- list(\n", paste(phase_entries, collapse = ",\n"), "\n)")
} else {
  phase_str <- paste0(
    "PHASE_MAP <- list(\n",
    '  # "01-Phase" = "Phase"    # \u2190 fill in your phase directories\n',
    ")"
  )
}

# Build NOTES_EXCLUDE string
if (length(NOTES_EXCLUDE) > 0) {
  exclude_str <- paste0('NOTES_EXCLUDE <- c("', paste(NOTES_EXCLUDE, collapse = '", "'), '")')
} else {
  exclude_str <- "NOTES_EXCLUDE <- c()"
}

new_config <- c(
  '# ── Config ──────────────────────────────────────────────────────────────────',
  paste0('BASE_DIR   <- "', TARGET_DIR, '"'),
  'HTML_FILE  <- file.path(BASE_DIR, "model-explorer.html")',
  'source(file.path(BASE_DIR, "dashboard", "validate-models.R"))',
  '',
  phase_str,
  '',
  'IGNORE_MODELS <- c()',
  '',
  'PLOT_DEFAULTS <- list(',
  '  lloq           = NA,',
  '  yLimitLower    = NA,',
  '  dataIsLog      = FALSE,',
  '  logY           = FALSE,',
  '  allowLogToggle = TRUE',
  ')',
  '',
  exclude_str,
  ''
)

gen_lines <- c(gen_lines[1:(config_start - 1)], new_config, gen_lines[(config_end + 1):length(gen_lines)])
writeLines(gen_lines, gen_path)
cat("\u2713 Patched config in", gen_path, "\n")
```

## 5. Next steps

After knitting this notebook:

1. **Organize your Monolix projects** under `r TARGET_DIR` following the layout above
2. **Edit** `generate-explorer-data.R` if you need to adjust `PHASE_MAP`,
   `PLOT_DEFAULTS`, or `IGNORE_MODELS` (config was pre-filled from above)
3. **Generate the dashboard data:**
    ```r
    source("dashboard/generate-explorer-data.R")
    ```
    or: `bash dashboard/update-dashboard.sh`
4. **Open** `model-explorer.html` in your browser

To refresh after re-running models, just re-run step 3.

## Dashboard features

| Tab | What it shows |
|-----|---------------|
| **Overview** | BICc/BIC/AIC ranking, convergence status, parameter counts |
| **Parameters** | Detailed estimates + RSE, Structure (FIXED/FITTED), Initial values |
| **Shrinkage** | η-shrinkage per random effect, color-coded by severity |
| **Comparison** | All models in the phase ranked by selected criterion |
| **Parm Plot** | Violin plots of individual parameter distributions (EBE modes) |
| **Ind. Fits** | Per-subject time-course fits with observed data overlay |
| **Diagnostics** | Observed vs IPRED / Pop Pred scatter plots |
